### 2020-05-18 docs/paper/LaTex/Paper.tex
Sorting files and folders again

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%   8888888b.                                         888      888
%   888   Y88b                                        888      888
%   888    888                                        888      888
%   888   d88P 888d888 .d88b.   8888b.  88888b.d88b.  88888b.  888  .d88b.
%   8888888P"  888P"  d8P  Y8b     "88b 888 "888 "88b 888 "88b 888 d8P  Y8b
%   888        888    88888888 .d888888 888  888  888 888  888 888 88888888
%   888        888    Y8b.     888  888 888  888  888 888 d88P 888 Y8b.
%   888        888     "Y8888  "Y888888 888  888  888 88888P"  888  "Y8888

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% vscode extension LaTex-Workshop uses these magic comments to 
% define which program will be used to build the project or bibliography
% I do need lualatex to use markdown package. Also, when I use lualatex
% in need to include `pdfencoding=auto` in hyperref setup.
% !TEX program = lualatex
%% !TEX options = -synctex=1 -interaction=nonstopmode -shell-scape "%DOC%"
% `-shell-scape` argument doesn't work in Windows 10 lualatex
% !TEX options = -synctex=1 -interaction=nonstopmode "%DOC%"
% !BIB program = biber

\documentclass[11pt,a4paper,dvipsnames,twoside]{article}

% Paper geometry
\usepackage[a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}
% \usepackage[showframe,a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}

% inputenc is ignored when I'm using lualatex as builder
% \usepackage[utf8]{inputenc}

% To produce a spanish automated output, for instance, TOC is called 'Índice' and
% References is called 'Referencias'
\usepackage[english]{babel}

% Modifier for space between lines (=\baselinestretch*\baselineskip)
% \renewcommand{\baselinestretch}{1.5}
% Modifies the space between notes in footnotes 
% IMPORTANT: just the same than paragraphs (parskip package below)
% \setlength{\footnotesep}{.9\baselineskip}

% It is better to use setspace package
\usepackage{setspace}
\setstretch{1.5}

% Modifies the space over footnote line
\addtolength{\skip\footins}{\baselineskip}

% To convert fonts already installed in the system and using them as main font in the document
% To check the names that fontspec admits use `otfinfo -i </path/to/font/file/even.ttf>`
% otfinfo does admit also ttf files! For instance:
% `otfinfo -i /usr/share/fonts/liberation/LiberationSerif-Regular.ttf`
\usepackage{fontspec}
\setmainfont{Liberation Serif}
\setsansfont{Liberation Sans}
\setmonofont{CourierPrimeCode-Regular} 
% \setmonofont{Liberation Mono}
% A different font for titles
\newfontfamily\headingfont[]{Liberation Sans}

% To allow custom headings
\usepackage[compact]{titlesec}
\titleformat*{\section}{\Large\headingfont\bfseries}
\titleformat*{\subsection}{\large\headingfont\bfseries}
\titleformat*{\subsubsection}{\normalsize\headingfont\bfseries}
% To control spacing for sections, I'm not using it, it just an example
% \titlespacing*{\section}{3em}{0em}{-2em}

% To change title font family
% \usepackage{titling}
% \renewcommand{\maketitlehooka}{\headingfont}

% To change TOC appearance; see 2.1 section in tocloft.pdf for info about 'titles' option
% it does not mention fancyhdr package but it do the work to flushright page numbers for 
% table of contents pages. see: https://tex.stackexchange.com/a/137301/83189
% \usepackage[titles]{tocloft}
\usepackage{tocloft}
\renewcommand{\cftsecfont}{\headingfont\bfseries}
\renewcommand{\cftsecpagefont}{\headingfont\bfseries}
\renewcommand{\cftsubsecfont}{\headingfont}
\renewcommand{\cftsubsecpagefont}{\headingfont}
\renewcommand{\cftsubsubsecfont}{\headingfont}
\renewcommand{\cftsubsubsecpagefont}{\headingfont}
\tocloftpagestyle{fancy} % However, I find better this way bc 'titles' option does not 
% produce a compact TOC, it does expand bc probably it uses the same spacing rules in 
% the TOC than in (subsub)(sub)sections titles at the rest of the text.
% -------------------
% Remove the TOC title (it is redundant if you have included a TOC section inside TOC)
\makeatletter
\renewcommand{\@cftmaketoctitle}{}
\makeatother
% But in the case you still want to print the TOC title, this other one modifies TOC title font
\renewcommand{\cfttoctitlefont}{\Large\headingfont\bfseries}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%   8888888b.                   888
%   888   Y88b                  888
%   888    888                  888
%   888   d88P 8888b.   .d8888b 888  888  8888b.   .d88b.   .d88b.  .d8888b
%   8888888P"     "88b d88P"    888 .88P     "88b d88P"88b d8P  Y8b 88K
%   888       .d888888 888      888888K  .d888888 888  888 88888888 "Y8888b.
%   888       888  888 Y88b.    888 "88b 888  888 Y88b 888 Y8b.          X88
%   888       "Y888888  "Y8888P 888  888 "Y888888  "Y88888  "Y8888   88888P'
%                                                      888
%                                                 Y8b d88P
%                                                  "Y88P"

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Basic packages (they don't need extra configurations)
\usepackage{% `esvect` will not work without texlive-fontsextra
  amsmath,
  amsfonts,
  amssymb,
  graphicx,
  multirow,
  tabularx,
  mathtools,
  caption,
  subcaption,
  color
}

% Extra colors
\usepackage[x11names,table]{xcolor}

% To cross references and metadata
\usepackage[]{hyperref}
\hypersetup{%
  pdfencoding=auto,
  pdftitle={Degree Final Project},
  pdfauthor={Javier Fernández // jfernandil@alumnos.unex.es},
  pdfsubject={},
  pdfkeywords={Project, Dendrometry, LoRa, Arduino, IoT},
  colorlinks = false,
  linkbordercolor = DarkGoldenrod2,
  urlbordercolor= DarkGoldenrod2,
  %linkcolor = black,
  %urlcolor  = black,
  breaklinks
}

% To make the frame around UEx seal
\usepackage{mdframed}

% To include References and TOC in the TOC
\usepackage[]{tocbibind}

% To avoid 'underfull box' warning caused because bibliography line-break
\usepackage{etoolbox}
\apptocmd{\thebibliography}{\raggedright}{}{}

% To use \begin{markdown}\end{markdown} environment and be able to type markdown mode
% IMPORTANT: do not use indent inside \begin{markdown}\end{markdown} environment in the code! 
% if you do lualatex is not translating the markdown code.
\usepackage{markdown}

% To get proper skips between paragraphs when you left a blank line in the code! (so I don't need to use `\\` 
% and I'm not getting warnings all the time) 
% IMPORTANT:
% here is an interesting article speaking about paragraphs an indentation, there are two types of paragraphs; 
% - classic: using indentation to distinguish between one and another
% - modern: dispenses with indents and instead distinguish paragraphs using a blank line between them
% Commenting/uncommenting the following package you can choose between two types.
% more info here: https://tex.stackexchange.com/a/196941/83189
% However this package also has options, which can be
% `indent` (to enable indentation)
% `skip=.5\baselineskip` (by default)
% `parfill` (to impose a minimum space at the end of the last line of a paragraph =30pt by default)
% \usepackage[skip=\baselineskip]{parskip}
% Another different option
% https://tex.stackexchange.com/a/371986/83189
\parskip=.75\baselineskip
\parindent=0pt

% To make quotes (I need renewcommand for mkbegdispquote and mkenddispquote to enable quote marks at the
% end and beginning to displayquote environment. More info: https://tex.stackexchange.com/questions/101809/how-to-force-csquotes-to-show-the-quotations-marks-in-block-quotes)
\usepackage[style=english]{csquotes}
\renewcommand{\mkbegdispquote}[2]{\openautoquote}
\renewcommand{\mkenddispquote}[2]{\closeautoquote#1#2}
\renewcommand{\mkcitation}[1]{ #1}

% To print units
\usepackage{siunitx}

% Enumerate lists
\usepackage{enumitem}
% To fix spaces between paragraphs, items and top or bottom fo enumerate/itemize environments I need the following
% see page 3 in https://osl.ugr.es/CTAN/macros/latex/contrib/enumitem/enumitem.pdf
\setlist[itemize]{%
  parsep=.75\baselineskip,
  topsep=0pt,
  partopsep=0pt,
  itemsep=0pt
}
\setlist[enumerate]{%
  parsep=.75\baselineskip,
  topsep=0pt,
  partopsep=0pt,
  itemsep=0pt
}

% To backrefs in the footnotes
\usepackage{footnotebackref}

% To improve headers an footers (I'm using to set page number at right)
\usepackage{fancyhdr}
% Turn on the style
\pagestyle{fancy}
\fancyhf{} % Start with clearing everything in the header and footer
\renewcommand{\headrulewidth}{0pt} % Removes the horizontal rule in header by default
\fancyfoot[R]{\thepage} % Set the right side of the footer to be the page number

% To insert json snippets
\usepackage{listings}
\renewcommand{\lstlistingname}{Ejemplo}

% \colorlet{punct}{red!60!black}
% \definecolor{background}{HTML}{EEEEEE}
% \definecolor{delim}{RGB}{20,105,176}
% \colorlet{numb}{magenta!60!black}
\definecolor{punct}{HTML}{0064d3}
\colorlet{background}{gray!20}
\definecolor{delim}{HTML}{ff0000}
\definecolor{numb}{HTML}{0f9d58}

\lstdefinelanguage{json}{
  % aboveskip=\baselineskip,
  % xleftmargin=2em,
  basicstyle=\scriptsize\ttfamily,
  numbers=left,
  numberstyle=\scriptsize,
  stepnumber=1,
  numbersep=8pt,
  showstringspaces=false,
  breaklines=true,
  frame=lines,
  backgroundcolor=\color{background},
  literate=
   *{0}{{{\color{numb}0}}}{1}
    {1}{{{\color{numb}1}}}{1}
    {2}{{{\color{numb}2}}}{1}
    {3}{{{\color{numb}3}}}{1}
    {4}{{{\color{numb}4}}}{1}
    {5}{{{\color{numb}5}}}{1}
    {6}{{{\color{numb}6}}}{1}
    {7}{{{\color{numb}7}}}{1}
    {8}{{{\color{numb}8}}}{1}
    {9}{{{\color{numb}9}}}{1}
    {:}{{{\color{punct}{:}}}}{1}
    {,}{{{\color{punct}{,}}}}{1}
    {"}{{{\color{punct}{"}}}}{1}
    {\{}{{{\color{delim}{\{}}}}{1}
    {\}}{{{\color{delim}{\}}}}}{1}
    {[}{{{\color{delim}{[}}}}{1}
    {]}{{{\color{delim}{]}}}}{1},
}

\newcommand\YAMLcolonstyle{\color{red}\mdseries}
\newcommand\YAMLkeystyle{\color{black}\mdseries}
\newcommand\YAMLvaluestyle{\color{blue}\mdseries}
% here is a macro expanding to the name of the language
% (handy if you decide to change it further down the road)
\lstdefinelanguage{yaml}{
  keywords={true,false,null,y,n},
  keywordstyle=\color{darkgray}\bfseries,
  basicstyle=\YAMLkeystyle\scriptsize\ttfamily,
  numbers=left,
  numberstyle=\scriptsize,
  numbersep=8pt,
  breaklines=true,
  frame=lines,
  backgroundcolor=\color{background},
  sensitive=false,
  comment=[l]{\#},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{purple}\ttfamily,
  stringstyle=\YAMLvaluestyle\ttfamily,
  moredelim=[l][\color{orange}]{\&},
  moredelim=[l][\color{magenta}]{*},
  moredelim=**[il][\YAMLcolonstyle{:}\YAMLvaluestyle]{:},   % switch to value style at :
  morestring=[b]',
  morestring=[b]",
  literate = {---}{{\ProcessThreeDashes}}3
             {>}{{\textcolor{red}\textgreater}}1     
             {|}{{\textcolor{red}\textbar}}1 
             {\ -\ }{{\mdseries\ -\ }}3,
}

% To highlight text with automated line break or 'Overfull \hbox' messages
\usepackage{soul}
\soulregister\enquote7

% biblatex package
\usepackage[block=ragged,style=ieee]{biblatex}
\addbibresource{Paper.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%    .d8888b.                                                              888
%   d88P  Y88b                                                             888
%   888    888                                                             888
%   888         .d88b.  88888b.d88b.  88888b.d88b.   8888b.  88888b.   .d88888 .d8888b
%   888        d88""88b 888 "888 "88b 888 "888 "88b     "88b 888 "88b d88" 888 88K
%   888    888 888  888 888  888  888 888  888  888 .d888888 888  888 888  888 "Y8888b.
%   Y88b  d88P Y88..88P 888  888  888 888  888  888 888  888 888  888 Y88b 888      X88
%    "Y8888P"   "Y88P"  888  888  888 888  888  888 "Y888888 888  888  "Y88888  88888P'

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Provides the \doubt{} command
\newcommand{\doubt}[1] {\textbf{\color{red}#1}}

% Provides the \cmd{} command; using soul package and previous \letcolor
% definition for listings package. Read page 3 of soul.pdf or 
% https://tex.stackexchange.com/a/231647/83189
% I need that to get a correct and automated line break when the command I want
% to highlight is too long and is placed at the end of the line.
\sethlcolor{background}
\newcommand{\cmd}[1] {\hl{\texttt{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%   88888888888 d8b 888    888
%       888     Y8P 888    888
%       888         888    888
%       888     888 888888 888  .d88b.
%       888     888 888    888 d8P  Y8b
%       888     888 888    888 88888888
%       888     888 Y88b.  888 Y8b.
%       888     888  "Y888 888  "Y8888

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\headingfont%
  University of Extremadura\\
  Faculty of Science\\[1cm]
  \textbf{%
  	Physics degree\\
	  Degree Final Project\\[3cm]
	  \flushleft{\large{About implementation of LoRa-based wireless monitoring systems for tree species dendrometry.}}\\[3cm]
  }
}

% I've used that \parbox to flush right the author, but I must have use this package https://osl.ugr.es/CTAN/macros/latex/contrib/titling/titling.pdf
\author{%
  \parbox{.9\textwidth}{
    \begin{flushright}
      Javier Fernández Aparicio\\
      % $\blacktriangledown$\\
      \href{mailto:jfernandil@alumnos.unex.es}{jfernandil@alumnos.unex.es}\\[\dimexpr\baselineskip + 1cm]
      July 2020
    \end{flushright}
  }
}
\date{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%   888888b.                     d8b               8888888b.
%   888  "88b                    Y8P               888  "Y88b
%   888  .88P                                      888    888
%   8888888K.   .d88b.   .d88b.  888 88888b.       888    888  .d88b.   .d8888b
%   888  "Y88b d8P  Y8b d88P"88b 888 888 "88b      888    888 d88""88b d88P"
%   888    888 88888888 888  888 888 888  888      888    888 888  888 888
%   888   d88P Y8b.     Y88b 888 888 888  888      888  .d88P Y88..88P Y88b.
%   8888888P"   "Y8888   "Y88888 888 888  888      8888888P"   "Y88P"   "Y8888P
%                            888
%                       Y8b d88P
%                        "Y88P"

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% To manage hyphenation (words break at end of the line), in this case is to prevent "Tellink" breakage.
\hyphenation{Tellink}

\begin{document}
\noindent % to prevent the indentation before first minipage and be able to expand minipages
\begin{minipage}[t]{.28\textwidth}
    \centering
      \begin{mdframed}[innerbottommargin=490pt, innertopmargin=40pt, linewidth=1pt]
        \includegraphics[width=\textwidth]{../../pictures/Seal/marca-uex-2-color.png}
      \end{mdframed}
\end{minipage}
\hskip 30pt minus 5pt % It is a dynamic space (glue) that starts on 30pt and is 5pt shrinkable
\begin{minipage}[t]{.6\textwidth}
  \centering
  \vspace{1cm}
  \maketitle
\end{minipage}
\thispagestyle{empty}
\newpage
\begingroup
  \thispagestyle{empty}
  \null%
\endgroup
\newpage
% \pagenumbering{arabic}% Arabic page numbers (and reset to 1)
\setcounter{page}{1}
\tableofcontents

\newpage

\section{Abstract}
This document gives a detailed description of this project, which is focused on researching possibles low-cost alternatives for wireless dendrometry systems. Currently there exist a lot of expensive and professional systems in the market, that's because this project is intended to reduce costs and increase the versatility, scalability and accessibility.

In order to reach these objetives the project will be supported over free software such as FIWARE\cite{Fiware} or free hardware such as Arduino\cite{Arduino} and RaspberryPi\cite{Raspberrypi}.

\printbibliography
\end{document}